---
title: "ElasticNet"
author: "Hunter Manhart"
date: "November 11, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Notebook for running ridge-regression and lasso on the SAheart dataset, using cross-validation

We need the glmnet package:
```{r, eval=FALSE}
install.packages("glmnet", repos = "http://cran.us.r-project.org")
```

Also, load the libraries:
```{r, message=FALSE, warning=FALSE}
library(glmnet)
library(ElemStatLearn)
```

Make famhist numeric
```{r}
SAh=as.matrix(SAheart[,-5])
famhist <- as.numeric(SAheart[,5])
SAh=cbind(SAh[,1:4],famhist,SAh[,5:9])
```

Use CHD, LDL, and adiposity as output variables
```{r}
XCHD = SAh[,-10]
yCHD = SAh[,10]

XLDL = SAh[,-3]
yLDL = SAh[,3]

XADI = SAh[,-4]
yADI = SAh[,4]
```

Preform ridge-regression
```{r}
ridgeCHD = glmnet(XCHD,yCHD, alpha=0)
ridgeLDL = glmnet(XLDL,yLDL, alpha=0)
ridgeADI = glmnet(XADI,yADI, alpha=0)
```

Plot
```{r}
plot(ridgeCHD)
vnat=coef(ridgeCHD)
vnat=vnat[-1,ncol(vnat)] # remove the intercept, and get the coefficients at the end of the path
axis(4, at=vnat,line=-.5,label=colnames(XCHD),las=1,tick=FALSE, cex.axis=0.5)
```

```{r}
plot(ridgeLDL)
vnat=coef(ridgeLDL)
vnat=vnat[-1,ncol(vnat)] # remove the intercept, and get the coefficients at the end of the path
axis(4, at=vnat,line=-.5,label=colnames(XLDL),las=1,tick=FALSE, cex.axis=0.5)
```

Had to print this on lambda axis because the graph strangely curves back on itself in L1
```{r}
plot(ridgeADI, xvar='lambda')
vnat=coef(ridgeADI)
vnat=vnat[-1,ncol(vnat)] # remove the intercept, and get the coefficients at the end of the path
axis(4, at=vnat,line=-.5,label=colnames(XADI),las=1,tick=FALSE, cex.axis=0.5)
```


Preform lasso
```{r}
lassoCHD = glmnet(XCHD,yCHD, alpha=1)
lassoLDL = glmnet(XLDL,yLDL, alpha=1)
lassoADI = glmnet(XADI,yADI, alpha=1)
```

Plot
```{r}
plot(lassoCHD)
vnat=coef(lassoCHD)
vnat=vnat[-1,ncol(vnat)] # remove the intercept, and get the coefficients at the end of the path
axis(4, at=vnat,line=-.5,label=colnames(XCHD),las=1,tick=FALSE, cex.axis=0.5)
```

```{r}
plot(lassoLDL)
vnat=coef(lassoLDL)
vnat=vnat[-1,ncol(vnat)] # remove the intercept, and get the coefficients at the end of the path
axis(4, at=vnat,line=-.5,label=colnames(XLDL),las=1,tick=FALSE, cex.axis=0.5)
```

```{r}
plot(lassoADI)
vnat=coef(lassoADI)
vnat=vnat[-1,ncol(vnat)] # remove the intercept, and get the coefficients at the end of the path
axis(4, at=vnat,line=-.5,label=colnames(XADI),las=1,tick=FALSE, cex.axis=0.5)
```



```{r}
cvfit = cv.glmnet(XADI,yADI, alpha=1)
plot(cvfit)
```
```{r}
cvfit$lambda.min
```


Age, LDL, and obesity appear to be the most important factors for adiposity (based on the adiposity lasso)
```{r}
Xless = cbind(SAh[,3], SAh[,7], SAh[,9])
```

We now have a similarly small mean-squared error using only 3 of the 9 input variables
```{r}
cvfit = cv.glmnet(Xless,yADI, alpha=1)
plot(cvfit)
```
```{r}
cvfit$lambda.min
```